---
title: 
author: "Yizhuo Liu, Leo Cheng, Haobo Ren"
date: "October-02-2024"
output: pdf_document
subtitle: 
date-format: long
abstract: 
format: pdf
number-sections: true
bibliography: references.bib
thanks: 'Code and data are available at: https://github.com/HaoboRrrr/FIFA20_Player_Potential_Rating_Analysis/tree/main'
---
```{r}
#| include: false
#| warning: false
#| message: false

library(tidyverse)
library(knitr)
```


# Contributions


# 1 Introduction


# 2 Data Description
The data of players in FIFA20 was originally found on Kaggle https://www.kaggle.com/datasets/stefanoleone992/fifa-20-complete-player-dataset?select=players_20.csv, it was scraped from the publicly available website using python script https://sofifa.com. \newline

## Responsive Varible
The response variable used in this research was "overall", it is the overall attribute of one player. This variable measures the players overall rating in the game, ranging from 0 to 100.\newline

Table 1: Statistics of overall \newline
```{r}
#| fig-cap: Statistical analysis of "overall'
#| echo: false
#| message: false
#| warning: false
#| information: false

data <- read.csv("../data/cleaned_data/cleaned_data.csv")
summary_stats <- summary(data$overall)
summary_stats
```
The mean of the overall is 66.43 \newline


## Linear Assumptions

The responsive variable "overall" is the overall rating of one player, we believe the overall rating of a player is largely influenced by their attributes like age, wages, pace score, shooting score and so on. When the player gets higher score in these attribute, their overall rating should also be higher. We think the changes in predictors should result in proportional changes in the response variable.

```{R}
#| echo: false
#| message: false
#| warning: false
#| information: false
#| fig-cap: Response against predictor test
plot(data[, c(2,3,4,5,6,7)])
```
The response variable has a linear relationship with the predictor variables, meaning that as the predictors change, the response variable changes in a consistent, straight-line fashion.\newline

```{R}
#| echo: false
#| message: false
#| warning: false
#| information: false
#| fig-cap: Residual against fitted value test
fit_rating = lm(overall ~ age + wage_eur + international_reputation +passing + physic, data = data)
fitted_value = fitted(fit_rating)
residual = resid(fit_rating)
plot(fitted_value, residual, xlim=c(50, 100))
```
The residuals are independent, meaning that the error in one observation doesn’t affect the error in another

```{R}
#| echo: false
#| message: false
#| warning: false
#| information: false
#| fig-cap: Response against fitted value test


plot(fitted_value,data$overall, xlim=c(50,100))


```
points closely follow the 45-degree line (i.e., observed ≈ predicted), the model is likely performing well.

# Preliminary Results


\newpage

# References